---
title: "COMPASScovariate Test Simulations"
author: "Tyler Schappe"
date: "2022-11-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

#Source for now but will want to call using COMPASS.R eventually
source("~/project_repos/COMPASScovariate/R/COMPASS-covariate.R")
source("~/project_repos/COMPASScovariate/R/updatebeta.R")
source("~/project_repos/COMPASScovariate/R/utils.R")

library(foreach); library(doParallel)
```

## Notes

Alternative sim method (prioritize this one):

1. Generate X cov
2. Specify true betas (non-zero simulated from Normal)
3. Run logistic regression to compute p-hat
4. Using p-hat, generate Gamma matrix via rbinom()
5. Generate pu, which is props for unstim
6. Generate counts via multinomial
7. Conditioned on gamma and pu, generate ps
8. Generate ns (counts) from ps (prob vector)


Using real data (COMPASS vignette):

1. Run COMPASS on real data
2. Get Gamma matrix
3. Get ns and nu (counts)
4. Generate X 
5. Run logistic regression (glmnet with lasso) between gamma and X. Estimated beta will be considered "true" beta. 

Note on iterations:

Keep only the last X iterations and use the first ones to warm up chains.

## Initial Simulation Method by Lynn

Generate combinations of simulation parameters
```{r}
sim.pars <- expand.grid(I = c(200, 30),
                        K = c(10, 16),
                        p = c(5)
            )
```

Run simulations corresponding to parameters above in parallel
```{r}
cl <- makeCluster(4)
registerDoParallel(cl)

sim.out.list <- foreach(s = 1:nrow(sim.pars), .packages = c("COMPASS")) %dopar% {

  I <- sim.pars[s,]$I ## sample size i.e., #subjects; I = 30
  K <- sim.pars[s,]$K ## number of cell categories: K = 16
  p <- sim.pars[s,]$p # number of covariates # larger p
  K1 = K-1
  
  set.seed(123)
  gamma <- matrix(0, nrow = I, ncol =  K) ## 0/1 response
  
  X <- matrix(rnorm(I*p), ncol = p, nrow = I) # covariates
  mu_true = rnorm(K, 0, 1) # the true intercept
  alpha_true = matrix(rbinom(p*K1, 1, 0.7), nrow = K1) # variable selection indicators
  
  B_true = matrix(rnorm(K1*p, 0, 5), nrow = K1)
  Balpha = B_true*alpha_true # coefficients with zeros
  
  
  for(k in 1:K1){
    z = mu_true[k] + X%*%(Balpha[k,]) # linear combination with a bias
    pr = 1/(1+exp(-z))   # pass through an inv-logit function
    gamma[,k] = unlist(lapply(pr, function(x) rbinom(1,1, x)))
  }
  
  gamma[,K] = rbinom(I,1,0.5)
  gamma_true = gamma
  
  pu_true = MCMCpack::rdirichlet(I, rep(10, K1))
  n_u = array(0, dim = c(I, K))
  Nu = ceiling(rnorm(I,20000, 10))
  for(i in 1:I){
    n_u[i,1:K1] = rmultinom(1, 5000, pu_true[i,])
  }
  n_u[,K] = Nu - rowSums(n_u)
  ps_true = pu_true
  
  #If gamma is 0, the pu = ps
  #
  for(i in 1:I){
    l1 = which(gamma[i,1:K1]==1)
    l0 = which(gamma[i,1:K1]==0)
    ps_true[i,l1] = MCMCpack::rdirichlet(1, rep(20, length(l1)))*sum(pu_true[i,l1])
  }
  
  n_s = array(0, dim = c(I, K))
  Ns = ceiling(rnorm(I,30000, 10))
  #Generating counts using true proportions
  for(i in 1:I){
    n_s[i,1:K1] = rmultinom(1, 6000, ps_true[i,])
  }
  n_s[,K] = Ns - rowSums(n_s)
  
  
  N = 10000
  K1 <- K - 1L
  p1 = p + 1L #
  beta =  array(0, dim=c(K1,p1, N)) #K-1 subsets, p features + 1 intercept
  nu =  invgamma::rinvgamma(p, shape = 0.5, rate = 1)
  u = invgamma::rinvgamma(K1, shape = 0.5, rate = 1)
  lambda2 = invgamma::rinvgamma(p, shape = 0.5, rate = 1/nu)
  tau2 = invgamma::rinvgamma(K1, shape = 0.5, rate = 1/u)
  W =  matrix(sapply(1:(I*K1), function(x) pgdraw::pgdraw(1,0)), nrow = I)
  sig2k = rep(10, K1) # var for normal prior of intercept
  X1 = cbind(1,X)
  
  
  
  fit_cov = .COMPASS.covariate(n_s, n_u, 1000, 2, X)
  for(tt in 2:N){
    res4 = updatebeta(gamma, X1, nu, u, lambda2, tau2, W, K1,p1, sig2k)
    W = res4$W
    u = res4$u
    nu = res4$nu
    tau2 = res4$tau2
    lambda2 = res4$lambda2
    beta[,,tt] = res4$beta
  }
  
  #Output the COMPASS fit object
  return(fit_cov)
}

#Assign names to the output list with param values
names(sim.out.list) <- apply(sim.pars, MARGIN = 1, FUN = function(x) {
  #Combine the list with paste
  do.call(paste,
          c(
            #Loop over the columns of sim.pars and paste the colnames with the value
            lapply(1:ncol(sim.pars), FUN = function(c) {
              paste(colnames(sim.pars)[c],
                    x[c],
                    sep = ":"
                    )
            }),
          sep = "_")
          )
})

stopCluster(cl)
```

